---
title: "workshop_7"
author: "Kyle Beall"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##loading libraries
```{r}
library(tidyverse)
library(tidymodels)
library(skimr)
library(janitor)
```

##to remove old environment data
```{r}
rm(list=ls())
```


##loading the data
```{r}
volcano_raw<-read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv')
```


```{r}
skim(volcano_raw)

volcano_raw%>%filter(last_eruption_year != "Unknown")
```

##we want to predict the volcano type. First, we want to see how many types we have. 
```{r}
volcano_raw%>%count(primary_volcano_type, sort = TRUE)

##there are 26 types of volcanoes. Now we want to go through the categories and see if we can narrow down the feild. 
```

##cleaning the dataset category names and making factors
```{r}
volcano_df <- volcano_raw %>%
  transmute(volcano_type = 
              case_when(str_detect(primary_volcano_type, "Stratovolcano") ~ "Stratovolcano",
                                     str_detect(primary_volcano_type, "Shield") ~ "Shield",
                                     TRUE ~ "Other"),
            volcano_number, latitude, longitude, elevation,
            tectonic_settings, major_rock_1) %>%
  mutate_if(is.character, factor)

volcano_df%>%count(volcano_type)
```

##splitting the data (partitioning)
```{r}
set.seed(123)
volcano_df_split<-initial_split(volcano_df)
volcano_df_train<-training(volcano_df_split)
volcano_df_test<-testing(volcano_df_split)

volcano_df_train%>%count(volcano_type, sort = TRUE)
volcano_df_test%>%count(volcano_type, sort = TRUE)
##there are not many shield examples in our testing set 
```

##splitting the data with bootstrapping to ensure enough data in all sets
##use this when data is unbalanced
##more accurate, but requires more cpu memory and processing time
```{r}
volcano_boot<-bootstraps(volcano_df)
```


##building the recipe
##use 'update_role' to reclassify variable
```{r}
volcano_recipe<-recipe(volcano_type~., data = volcano_df)%>%
  update_role(volcano_number, new_role = "Id")%>%
  step_other(tectonic_settings)%>%
  step_other(major_rock_1)%>%
  step_dummy(tectonic_settings, major_rock_1) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())
##always include 'step_zv' as it doesn't hurt

volcano_prep<-prep(volcano_recipe)
```

##extract data from the recipe
```{r}
juice(volcano_prep)
```
##if you need to apply the recipe to new data set, you need to use 'bake()'


##modelling
```{r}
install.packages("ranger")
library(ranger)


##need to choose number of trees 
rf_spec <- rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger")
```


